\section{Materials and Procedures}
%Provides a detailed set of instructions for implementing the method, including all materials required and all procedures followed. Describes methods used in evaluation and intercalibration or intercomparison studies. Presents data sources, data extraction protocols and analytical methods used in reviews and metaanalyses.

%Whenever possible and appropriate, manuscripts must present complete instructions for the recommended procedure, analogous to a good cookbook or an easy-to-use laboratory manual. Descriptions of equipment and apparatus must provide a similar level of detail regarding the contruction and operation of the device. Evaluation and intercalibration or intercomparison studies may refer to published descriptions of existing methods, but should describe in reproducible detail how the present study was conducted. Metaanalytical studies should provide details regarding data sources and extraction, and analytical methods used.
Our approach for systematic, full coverage underwater photographic surveys is to use a line wound around a drum as a guide. Unwinding the line under tension constrains motion as to generate spiral pattern. The curve traced by the tip of the line corresponds to the involute of a circle (i.e., the cross section of the drum), with constant separation distance between revolutions corresponding to the perimeter of the circle. A more cumbersome version of this was attempted in 2007 in the expedition reported in \cite{Camilli_2007} with the line coiled on the imaging platform and the diver having to let out the trackline spacing after completing each revolution, resulting in concentric circles. More recently, a `minute mosaic' \cite{gintert2012third} uses a rebar pin as a visual reference for a diver to complete three revolutions with increasing radius while collecting imagery for a 2D mosaic. Since it depends purely on the diver's assessment of distance to the pin, the areas covered varied from 19 to 44 m^{2}. This is also likely to result in variable image overlap between revolutions, affecting the quality of the composite and potentially limiting its value for systematic repeat surveys. 

Archimidean spirals, that resemble involutes of a circle, have been investigated in the reconstruction of surfaces in metrology \cite{Wieczorowski_2001} and in estimating patchy distributions \cite{Kalikhman_2006} though these uses are concerned with sparse sampling of the area of interest, while our focus is in an operationally simple way to achieve full photographic coverage of an area with limited field of view (and footprint) imaging. Our approach is practical for full coverage of \sim{100m^{2}} areas.

For a drum of radius $R$ and angle $\alpha$ along the circumference of the drum, the tip of the line is located at polar coordinates $r$, $\phi$:  
\begin{equation}
r = R\cdot\sqrt{1+\alpha^2} \\
\phi = \alpha - \arctan{\alpha} 
\end{equation}
The outer boundary of the involute, $r$, grows in a near linear fashion with the number of revolutions, while the path (or arc) length, $L$, is given by
\begin{equation}
L=\frac{R}{2}\cdot\alpha^2,
\end{equation}
growing quadratically with the number of revolutions. 
From a survey design point of view, the choice of drum diameter and length of line, $M$, determine the number of revolutions needed to unwind the line as well as the path length. Together with the swimming speed this determines the duration. The final radius $r_M$ is given by the hyptothenus of the right angle triangle formed by the drum radius and the line extended at right angles, $r_M = \sqrt{M^2 + R^2}$ and therefore the total angle swept to unwind the whole drum is 
\begin{equation}
\alpha_M = \sqrt{\left(\frac{r_M}{R}\right)^2 - 1}
\end{equation}
substituting for $r_M$ yields $\alpha_M = \frac{M}{R}$ and therefore $L_M = \frac{M^2}{2R}$. The time to complete a survey, $t_M$, at a swimming speed $s$ is then
\begin{equation}
t_M = \frac{L_M}{s} = \frac{M^2}{2R\cdot s}
\end{equation}
For a line 6m long, a drum of 0.15m in diameter and a swimming speed of 0.2 $m/s$, the path length of the spiral pattern is X m and will take Y seconds.

\subsection{Materials}
%"Materials" includes expendible and non-expendible supplies, equipment, and solutions. Materials should be listed as completely as possible, with careful attention to providing enough information to assist new users in adopting the method. Descriptions of needed equipment must note the essential features of the equipment in sufficient detail to allow users to obtain similar devices, or authors may refer to specific commercial products in lieu of a detailed description. The composition of any solutions must be stated explicitly (e.g. grams per liter of a specific compound).
We rely on three three major components for the data acquisition
\begin{enumerate}
\item stereo imaging package. An updated version of the one used in \cite{Henderson_2013} \cite{Camilli_2007}, integrated into a smaller package. We require properly calibrated stereo pair and sensor suite, as well as the pose of each sensor relative to the navigation center \cite{Johnson_Roberson_2013} \cite{Mahon_2008}.
\item drum and line. Typically a series of reef records are collect. Prior to starting, it is important to select the diameter of the drum and lenght of the line to determine the size of the patch and the image overlap from the spiral spacing. The resulting pattern will have the form shown in Figure \ref{fig:spiral}.

% photo of drum and line

% full equation and approximate equation for drum and line length


\item star picket or base
% photo of base / star picket

\end{enumerate}



    
  
  
  
  