\subsection{Procedures}
%A stepwise description of all procedures used, divided into sub-procedures as necessary for clarity. Sufficient detail must be provided that a reader may readily adopt and employ the method as described, at least under the conditions described.
%Limnology and Oceanography: Methods will not accept manuscripts that describe laboratory and field techniques, equipment, analyses, and other methods, in insufficient detail to be reproduced by others.

In the field using two people: a swimmer and an assistant:\begin{enumerate}
\item Select location, such that the spiral covers the area of interest.
\item Drive star picket at the center of the area of interest. 
\item Attach pole and drum to star picket.
\item Clip imaging package to line.
\item The swimmer pushes the imaging package forward while keeping tension on the line and the desired altitude. Continuing until the line has completely unwound from the drum. In case of rough conditions, the assistant stays by the pole tending the line (Figure \ref{fig:swimming}).
\item Once the survey is completed, the swimmer detaches the line from the imaging platform. The assistant coils the line and takes the pole and drum off the star picket. 
\item The swimmer then goes with the imaging platform back over the center of the survey, completing at least four `spokes'. This extra images are not strictly necessary but can help constrain the reconstruction by providing some high overlap imagery across the spiral path.
\end{enumerate}

The result of following this procedure is a set of overlapping images covering an area two orders of magnitude larger than the footprint of single image. In the case of our pose-instrumented package, we also have a noisy observations of GPS (if on the surface), depth and heading, pitch, roll while performing the spiral survey. The imagery and navigation logs are post processed using the ACFR pipeline \cite{Johnson_Roberson_2010} \cite{Mahon_2008} \cite{Johnson_Roberson_2013} to estimate camera poses and 3D composite meshes of the surveyed area. Multi-scale structural complexity indeces such as rugosity are then derived from those meshes \cite{Friedman_2012}.
For a single camera setup, SFM packages will provide a 3D reconstruction of the imagery. Additional steps are needed to recover scale and georeferenced position and orientation (7 unknown parameters). This can done by determining the $(x,y,z)$ position of two points and the depth $z$ of a third point on the survey area.
